---
title: "ASPM"
output: html_document
date: "2023-08-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The application of the Age-Structured Production Model (ASPM) approach as a diagnostic can help identify misspecification of the production function. If, in the absence of composition data (likelihood weighting set to 0), the ASPM fits well to the indices of abundance that have good contrast, then the production function is likely to drive the stock dynamics and indices will provide information about the absolute abundance ([Carvalho et al. 2017](https://www.sciencedirect.com/science/article/pii/S0165783616303113)). If there is not a good fit to the indices, then the catch data and production function alone cannot explain the trajectories of the indices of relative abundance. 

## Model inputs  
To run a stock synthesis model, 4 input files are required: starter, forecast, control, and data. The input files for the example model can be found within the `ss3diags` package and accessed as shown below. Also, if you do not have `r4ss` installed, you will need to install for this tutorial. 

```{r eval=FALSE}
install.packages("pak")
pak::pkg_install("r4ss/r4ss")
```

```{r message=FALSE, warning=FALSE}
library(r4ss) 

files_path <- system.file("extdata", package = "ss3diags")
dir_tmp  <- tempdir(check = TRUE)
dir_aspm <- file.path(dir_tmp, "profile")
dir.create(dir_aspm, showWarnings = FALSE, recursive = TRUE)
list.files(files_path)
file.copy(from = list.files(files_path, full.names = TRUE), to = dir_tmp)
```




You will need to make sure you have the [SS executable](https://github.com/nmfs-stock-synthesis/stock-synthesis) file either in your path or in the directory you are running the ASPM from (in this case `dir_aspm`). An easy way to get the latest release of stock synthesis is to use the `r4ss` function `get_ss3_exe()`.

```{r}
r4ss::get_ss3_exe(dir = dir_tmp, version = "v3.30.21")
```

## ASPM 
Once you have the 4 input files, you need to determine what components need to be turned off to run the ASPM. ASPM only depend on index of abundance and catch data, so any composition data, recruitment deviations, etc. need to be turned off. We provide an example that includes multiple types of data and recruitment deviations, however, the exact steps necessary for an individual model may vary depending on the complexity and components included. Therefore these steps may not be fully comprehensive for your model so be sure check what other components you may need to change.  
Below, we show how to use the `r4ss` functions to make all the necessary changes to the `control.ss` and `ss.par` files.  

### Generate files  

In order to get the `ss.par` file, you need to run the model once. This will also produce a new data and control file that we will use as the input files. We will run the model in `dir_tmp` then copy the new files over to `dir_aspm`. 


```{r}
r4ss::run(dir = dir_tmp, verbose = FALSE, exe = "ss3")
files <- c("data_echo.ss_new", "control.ss_new", "starter.ss", "forecast.ss", "ss.par", "ss3")
file.copy(from = file.path(dir_tmp, files), to = dir_aspm)
```

### Recruitment Deviations  

Set the recruitment devations in `ss.par` to 0. 

```{r}
par <- SS_readpar_3.30(parfile = file.path(dir_aspm, "ss.par"), 
                       datsource = file.path(dir_aspm, "data_echo.ss_new"),
                       ctlsource = file.path(dir_aspm, "control.ss_new"),
                       verbose = FALSE)
par$recdev1
par$recdev_forecast

par$recdev1[,"recdev"] <- 0

#Would run if forecasts recdevs were not already 0
#par$recdev_forecast[,"recdev"] <- 0
SS_writepar_3.30(parlist = par, 
                 outfile = file.path(dir_aspm, "ss.par"), 
                 overwrite = T, verbose = FALSE)
```

### Starter File 
Change the starter file to read from `ss.par` (`init_values_src = 1`) and to use the new data and control files. 
```{r}
starter <- SS_readstarter(file = file.path(dir_aspm, "starter.ss"), verbose = FALSE)
starter$datfile <- "data_echo.ss_new"
starter$ctlfile <- "control.ss_new"
starter$init_values_src <- 1
SS_writestarter(starter, 
                dir = dir_aspm, 
                overwrite = TRUE, 
                verbose = FALSE, 
                warn = FALSE)
```

### Control File 
Some things in the control file that may need to be changed, are phases for parameters such as: 

* length selectivity 
* age selectivity
* size selectivity  
* steepness
* sigmaR  
* recruitment deviations 

Additionally, the likelihood for some components may need to be turned off as well. You can do this using the lambda options in the control file. Some examples include: 

* length-composition data  
* age-composition data 
* size-composition data  
* initial F  

```{r}
control <- SS_readctl_3.30(file = file.path(dir_aspm, "control.ss_new"), 
                           datlist = file.path(dir_aspm, "data_echo.ss_new"),
                           verbose = FALSE)

control$SR_parms
# Would need to run if PHASES were positive for "steep" and "sigmaR"
#control$SR_parms[c(2,3),"PHASE"] <- c(-4,-4)
control$age_selex_parms
#would need to run if PHASES were positive for age selectivity
#control$age_selex_parms[,"PHASE"] <- control$age_selex_parms[,"PHASE"] * -1
control$size_selex_parms[,"PHASE"] <- control$size_selex_parms[,"PHASE"] * -1
control$recdev_early_phase <- -4
control$recdev_phase <- -2 

new_lambdas <- data.frame(like_comp = c(4,4,5,5,7,7,9,9,9,10),
                          fleet = c(1,2,1,2,1,2,1,2,3,1), 
                          phase = rep(1, 10),
                          value = rep(0, 10),
                          sizefreq_method = rep(1, 10))
new_lambdas
control$lambdas <- new_lambdas
control$N_lambdas <- nrow(new_lambdas)
SS_writectl_3.30(control, outfile = file.path(dir_aspm, "control.ss_new"),
                 overwrite = TRUE, verbose = FALSE)
```

### Run ASPM 

```{r}
r4ss::run(dir = dir_aspm, exe = "ss3", skipfinished = FALSE, verbose = FALSE)
```

### Visualize Results 

```{r}
mods <- SSgetoutput(dirvec = c(
  dir_tmp,
  dir_aspm
), verbose = FALSE)

mods_sum <- SSsummarize(mods, verbose = FALSE)

SSplotComparisons(mods_sum, legendlabels = c("Ref", "ASPM"),
                  subplots = c(2,8,14))
```

