---
title: "Hindcast Cross-Validation"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ss3diags)
```

Hindcast Cross-Validation is used to evaluate the prediction skill of a model by removing a portion of data, fitting the model to the subsequent time series, then projecting over the period that was removed and comparing the projected values to the observations that were omitted. Prediction skill is determined using the mean absolute scaled error (MASE) for input data (index of abundance, length composition, etc). In brief, the MASE score scales the mean absolute error (MAE) of forecasts (i.e., prediction residuals) to mean absolute error (MAE) of a naïve in-sample prediction, which is realized in the form of a simple 'persistence algorithm', i.e. index of abunduance will be the same next year as it is this year (see Eq. 3, p.5 in [Carvalho and Winker et al. 2021](https://www.sciencedirect.com/science/article/pii/S0165783621000874)). A MASE score > 1 indicates that the average model forecasts are worse than a random walk. Conversely, a MASE score < 1 indicates a level of prediction skill, for example, a MASE of 0.5 indicates that the model forecasts twice as accurately as a naïve baseline prediction; thus, the model has prediction skill.

Implementing the Hindcast Cross-Validation (HCxval) diagnostic in Stock Synthesis requires the same model outputs generated when running a retrospective analysis. Therefore, no additional step is needed for HCxval if conducted in conjunction with retrospective analysis (see how to conduct a [restrospective analysis](https://pifscstockassessments.github.io/ss3diags/articles/Retrospective-Analysis.html)). For this example we will use the same output created in the retrospective example.

```{r echo=FALSE, message=FALSE, results=FALSE}
library(r4ss) 

files_path <- system.file("extdata", package = "ss3diags")
dir_retro <- file.path(tempdir(check = TRUE), "retrospectives")
dir.create(dir_retro, showWarnings = FALSE)
file.copy(from = list.files(files_path, full.names = TRUE), to = dir_retro)
file.copy(from ="../../inst/extdata/ss", to = dir_retro)
r4ss::retro(dir = dir_retro, exe = "ss", years = 0:-5, verbose = FALSE)
```


```{r}
retro_mods <- r4ss::SSgetoutput(dirvec = file.path(dir_retro, "retrospectives", paste0("retro", seq(0,-5,by=-1))), verbose = F)
retroSummary <- r4ss::SSsummarize(retro_mods, verbose = F)
```

 
## Visualizing Output  

HCxval is implemented using `ss3diags::SSplotHCxval()`, which produces the novel HCxval diagnostic plot and computes the MASE scores for indices of abundance, mean lengths, or mean ages that have observations falling within the hindcast evaluation period.

Plotting HCxval for abundance indices requires the same step of summarizing the list of retrospective runs as for the retrospective analysis, which therefore only needs be done once. In this example we will use `retroSummary` created [here](https://pifscstockassessments.github.io/ss3diags/articles/Retrospective-Analysis.html). 

```{r}
r4ss::sspar(mfrow = c(1,2))
SSplotHCxval(retroSummary, subplots = "cpue", add = TRUE)
```

In the plots above, we see that for both fleets, the model has fairly good prediction skill (compared to a random walk) of index of abundance data. In the plots, the white points and white dashed line are the observed data that were included in the model with a truncated time series. The larger colored points are the observed data from each retrospective peel (i.e. data that was removed in that peel). The smaller colored point and dashed line show the model predicted value. The "Ref" line is the model run with the complete time series of data. The grey shaded areas represent the uncertainty of the data, with the darker portion indicating the portion that were included in the model and the lighter portion indicating which ones were removed and projected. The MASE scores displayed are the MASE and adjusted MASE in parentheses.  

### Composition Data  

To visualize the mean length and mean age, you will need to add an additional step. Because the forecast length- and age-composition information are stored as "ghost files" in the SS report.sso file, you need to use `ss3diags::SSretroComps()` to extract and summarize that information before creating the plots. 

```{r}
retroSummary_comps <- SSretroComps(retro_mods)

r4ss::sspar(mfrow = c(1,2))
hcl <- SSplotHCxval(retroSummary_comps, subplots = "len", add = TRUE, ylim = c(40,75))
```

### Customizing the Plot  

HCxval plots can be customized in many ways, some common features that you may want to specify are:  

* displaying only a specific fleet
* removing uncertainty intervals 
* adjusting the x- or y-axis
* not displaying the adjusted MASE value on the plot
  
Examples of each of these changes are shown below, incrementally making each adjustment. 

```{r}
r4ss::sspar(mfrow = c(2,2))
SSplotHCxval(retroSummary_comps, subplots = "len", add = TRUE, indexselect = 1)
SSplotHCxval(retroSummary_comps, subplots = "len", add = TRUE, 
             indexselect = 1, indexUncertainty = FALSE)
SSplotHCxval(retroSummary_comps, subplots = "len", add = TRUE, 
             indexselect = 1, indexUncertainty = FALSE, ylim = c(45,70))
SSplotHCxval(retroSummary_comps, subplots = "len", add = TRUE, 
             indexselect = 1, indexUncertainty = FALSE, ylim = c(45,70), show.mase.adj = F)


```

## Summary Table  
In addition to the HCxval plots, a summary statistics table can be produced using `ss3diags::SSmase()`. This table includes: 

* Index - the fleet  
* Season  
* MASE - MASE score 
* MAE.PR - mean absolute error of prediction residuals 
* MASE.base - naive prediction 
* MASE.adj - adjusted MASE score 
* n.eval - number of points prediction skill was calculated for 

Adjusted MASE gets invoked in cases where the inter-annual variation in the observed values is very small (default MAE < 0.1 for naive predictions log(y[t+1])-log(y[t])).
The reasoning is that prediction residuals must be already very accurate to fall below this threshold. The adjusted MASE essential keep the naive prediction MAE denominator of the MASE to a maximum. Below we show the effect of changing adjustment threshold from the default `MAE.base.adj = 0.1`


```{r}
mase1 = SSmase(retroSummary_comps, quant="len", MAE.base.adj = 0.1)
mase1
```

to a larger value `MAE.base.adj = 0.15`

```{r}
SSmase(retroSummary_comps, quant="len", MAE.base.adj = 0.15)
```

where `MASE` is the ratio of the mean absolute error of the prediction residuals `MAE.PR` to the residuals of the naive predictions `MAE.base`

```{r}
mase1$MAE.PR/mase1$MAE.base
mase1$MASE
```

and MASE.adj


```{r}
mase1$MAE.PR/pmax(mase1$MAE.base,0.1)
mase1$MASE.adj
```