---
title: "Jitter Analysis"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(ss3diags)
library(r4ss)
library(tidyverse)
```

Jitter analyses are commonly implemented in Stock Synthesis to ensure a model has reached global convergence. Jitter involves changing the parameter start values by a small increment and rerunning the model to see if that adjustment causes the model to converge at a lower likelihood. This can be useful for distinguishing if a model reached a local minimum or a global minimum. The number of jitter iterations should be anywhere between 50-100 to ensure a good spread of start values. If any of those runs has a lower likelihood than your current model, parameter start values should be adjusted to use those from the run with a lower likelihood. You can do this by adjusting the values in the control.ss file to match those in the ss.par_#_of_the_lower_likelihood_run. We provide the steps for running jitter analysis using `r4ss::jitter()` below. 

## Model inputs  
To run a Stock Synthesis model, four input files are required: starter.ss, forecast.ss, control.ss, and data.ss. The input files for the example model can be found within the `ss3diags` package and accessed as shown below. Also, if you do not have the `r4ss` package installed, you will need to install it for this tutorial. 

```{r eval=FALSE}
install_github("r4ss/r4ss")
```

```{r message=FALSE, warning=FALSE}
library(r4ss)

files_path <- system.file("extdata", package = "ss3diags")
dir_jitter <- file.path(tempdir(check = TRUE), "jitter")
dir.create(dir_jitter, showWarnings = FALSE, recursive = TRUE)
list.files(files_path)
file.copy(from = list.files(files_path, full.names = TRUE), to = dir_jitter)
```

You will need to make sure you have the [SS executable](https://github.com/nmfs-stock-synthesis/stock-synthesis) file either in your path or in the directory you are running the retrospective from (in this case `dir_jitter`). An easy way to get the latest release of stock synthesis is to use the `r4ss` function `get_ss3_exe()`.

```{r}
r4ss::get_ss3_exe(dir = dir_jitter)
```
 
We will run the model in `dir_jitter` first to produce the necessary output files. It is recommended to do jitter runs in a subdirectory of your model run. This will keep all of the output files separate from other diagnostic tests you may run. 
```{r}
r4ss::run(dir = dir_jitter, exe = "ss3", verbose = FALSE)
```

## Jitter 
For this example, we will run 50 jitters. The `jitter()` function automates the entire process so you only need to give it a few arguments and it will run and produce the total likelihoods for each run. Full documentation of the `jitter()` function can be found at the [r4ss website](https://r4ss.github.io/r4ss/reference/jitter.html).

```{r run-jitter}
Njitter <- 50
jit.likes <- r4ss::jitter(
  dir = dir_jitter,
  Njitter = Njitter,
  init_values_src = 1,
  jitter_fraction = 0.1,
  exe = "ss3",
  printlikes = FALSE,
  verbose = FALSE
)
```

To analyze the output of all 50 runs, use `r4ss::SSgetoutput()` and `r4ss::SSsummarize()` as shown below. 

```{r }
jit_mods <- SSgetoutput(
  keyvec = 0:Njitter, # 0 to include reference run (Report0.sso)
  getcomp = FALSE,
  dirvec = dir_jitter,
  getcovar = FALSE,
  verbose = FALSE
)
jit_summary <- SSsummarize(jit_mods, verbose = FALSE)
```

Some key sections you may want to check and compare across models are, likelihoods, derived quantities, and estimated parameters. 
```{r }
head(jit_summary$likelihoods)
head(jit_summary$quants)
head(jit_summary$pars)
```

You may also want to check that the models converged. To do this you can check the maximum gradients to make sure they are all small (< 0.0001). You can also compare the estimated spawning biomass, if they are extreme values (+2x or <0.5x the base spawning biomass) this suggests the model didn't converge.

```{r }
# Maximum gradient
converged_grad <- which(jit_summary$maxgrad < 0.0001)

converged_ssb <- jit_summary$SpawnBio |>
  dplyr::mutate(across(c(1:(Njitter + 1)),
    .fns = ~ . / replist0
  )) |> # for each column, divide SSB by SSB from the reference run (replist)
  dplyr::select(-Label) |>
  tidyr::pivot_longer(col = c(1:(Njitter + 1)), names_to = "jitter", values_to = "SSB") |>
  tidyr::pivot_wider(names_from = Yr, values_from = SSB) |>
  dplyr::mutate(rownumber = seq(1, dplyr::n())) |>
  tibble::column_to_rownames("rownumber") |>
  dplyr::select(-jitter) |>
  dplyr:: filter(across(everything(), ~ . >= 0.5 & . <= 2)) |> # keep only rows where SSB is a reasonable value
  tibble::rownames_to_column("rownumber") |>
  dplyr::select(rownumber) |>
  dplyr::pull(rownumber)
converged_mods <- intersect(converged_grad, as.numeric(converged_ssb)) # getting which models are in both groups
converged_jitters <- jit_mods[converged_grad]
converged_sum <- SSsummarize(converged_jitters, verbose = FALSE)
```


## Visualizing Output
To compare the likelihoods of all runs, we plot them as shown below. There are no built in functions (as of writing this vignette) in `r4ss` or `ss3diags` to generate a likelihood plot, therefore we provide code in the tidyverse syntax (using `ggplot2`) to visualize the results. 
```{r }
converged_sum$likelihoods |>
  filter(str_detect(Label, "TOTAL")) |>
  select(-Label) |>
  pivot_longer(cols = everything(), names_to = "jitter", values_to = "likelihood") |>
  separate(jitter, into = c("replist", "jitter"), sep = "(?<=[A-Za-z])(?=[0-9])") |>
  mutate(jitter = as.numeric(jitter)) |>
  ggplot(aes(x = jitter, y = likelihood)) +
  geom_point(size = 3) +
  geom_hline(aes(yintercept = likelihood[1]), color = "red") +
  theme_classic() +
  labs(
    y = "Total Likelihood",
    x = "Jitter runs at a converged solution"
  )
```

The figure plots the total likelihood from each jitter run and the red line indicates the total likelihood value from the reference run. If there are any runs with points below the red line, use the converged parameter values from the run with the lowest likelihood value as starting values in your base model. 
We can also use `r4ss::SSplotComparisons()` to compare the spawning biomass trajectories between jitter runs to see the impact of different parameter values on the estimated quantities. 
```{r }
SSplotComparisons(converged_sum,
  subplots = 2,
  ylimAdj = 1,
  new = FALSE
)
```

