---
title: "Retrospective Analysis"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ss3diags)
```

Retrospective analysis is commonly used to check the consistency of model estimates such as spawning stock biomass (SSB) and fishing mortality (F) as the model is updated with new data in retrospect. The retrospective analysis involves sequentially removing observations from the terminal year (i.e., peels), fitting the model to the truncated series, and then comparing the relative difference between model estimates from the full-time series with the truncated time-series. To implement a retrospective analysis with stock synthesis the `r4ss` package provides the `retro()` function. Here we provide a step-by-step example of how to run and analyze a retrospective analysis using a simple, example SS model.   

## Model inputs  
To run a stock synthesis model, 4 input files are required: starter, forecast, control, and data. The input files for the example model can be found within the `ss3diags` package and accessed as shown below. Also, if you do not have `r4ss` installed, you will need to install for this tutorial. 

```{r eval=FALSE}
install.packages("pak")
pak::pkg_install("r4ss/r4ss")
```

```{r message=FALSE, warning=FALSE}
library(r4ss) 

files_path <- system.file("extdata", package = "ss3diags")
dir_retro <- file.path(tempdir(check = TRUE), "retrospectives")
dir.create(dir_retro, showWarnings = FALSE)
list.files(files_path)
file.copy(from = list.files(files_path, full.names = TRUE), to = dir_retro)

```


You will need to make sure you have the [SS executable](https://github.com/nmfs-stock-synthesis/stock-synthesis) file either in your path or in the directory you are running the retrospective from (in this case `dir_retro`). An easy way to get the latest release of stock synthesis is to use the `r4ss` function `get_ss3_exe()`.

```{r}
r4ss::get_ss3_exe(dir = dir_retro, version = "v3.30.21")
```

## Retrospective Analysis
Once you have the 4 input files and SS executable, you can run retrospective analysis as shown below. We are running it for 5 1-year peels, so with each run, *n* years of data are removed from the reference model and the model is re-run for a total of 5 times (i.e. peel 1 removes the last year of data, peel 2 removes the last 2 years of data, etc.) . The number of year peels can be adjusted with the `years` argument. If the SS executable file you are using is named something other than "ss3" (e.g. ss_opt_win.exe), you will need to specify this with the argument `exe = "ss_opt_win"`. Full documentation of the `retro()` function can be found on the [r4ss website](https://r4ss.github.io/r4ss/reference/retro.html).

```{r message=FALSE}
r4ss::retro(dir = dir_retro, exe = "ss3", years = 0:-5, verbose = FALSE)

```
## Visualizing Output 

To visualize the output and inspect for any patterns or biases, you need to load the report files into R and can use the `SSplotRetro()` function from `ss3diags`. The easiest way to load multiple report files is using `r4ss::SSgetoutput()` and `r4ss::SSsummarize()` functions. The default sub-directories for each peel, 0 to 5, are labeled `retro0` to `retro-5`.

```{r get_retro_files, message=FALSE, warning=FALSE}
retro_mods <- r4ss::SSgetoutput(dirvec = file.path(dir_retro, "retrospectives", paste0("retro", seq(0,-5,by=-1))), verbose = F)
retroSummary <- r4ss::SSsummarize(retro_mods, verbose = F)
SSplotRetro(retroSummary, subplots = "SSB", add = TRUE)

```
The default settings plot the spawning stock biomass time series for each peel, with the reference run (e.g. model with no years removed) as the "Ref" line and each successive peel as colored lines labeled by their end year. The solid line ends at the end year and the dashed line to the point shows the 1 year forecast. Displaying the projected SSB can help assess forecast bias. Note, forecasts are done automatically when using `r4ss::retro()` and are based on the settings in forecast.ss. The grey shaded area represents the 95% confidence intervals of uncertainty around the spawning biomass time series. Displayed in the center of the plot is the combined Mohn's $\rho$ for all retrospective runs, and in parentheses is the forecast Mohn's $\rho$. 

### Customizing the Plot  

Retro plots can be customized in many ways, some common features that you may want to specify are:  

* removing uncertainty intervals 
* adjusting the years shown on the x-axis
* turning off the 1-year ahead forecasting
* not displaying the combined $\rho$ value on the plot
  
Examples of each of these changes are shown below, incrementally making each adjustment. 
```{r message=FALSE, warning=FALSE}
r4ss::sspar(mfrow = c(2, 2), plot.cex = 0.8)
retro1 <- SSplotRetro(retroSummary, subplots = "SSB", add = TRUE, uncertainty = FALSE)
retro2 <- SSplotRetro(retroSummary, subplots = "SSB", add = TRUE,  uncertainty = FALSE, xlim = c(2015, 2022))
retro3 <- SSplotRetro(retroSummary, subplots = "SSB", add = TRUE,  uncertainty = FALSE, xlim = c(2015, 2022), forecast = FALSE)
retro4 <- SSplotRetro(retroSummary, subplots = "SSB", add = TRUE,  uncertainty = FALSE, xlim = c(2015, 2022), forecast = FALSE, showrho = FALSE, forecastrho = FALSE)

```
Additionally, the fishing mortality can be plotted instead of spawning biomass by replacing `subplots = "SSB"` with `subplots = "F"`

### Summary Table  

In addition to the retrospective plots, a summary statistics table can be produced using `SShcbias()`. This table includes 

* type of estimate (SSB or F)
* the year removed or "peel"
* Mohn's $\rho$ 
* forecast bias 

by year and overall ("Combined"). Mohn's $\rho$ is a measure of the severity of bias in the retrospective patterns and the forecast bias is an estimate of bias in the forecasted quantities when years of data were removed. The rule of thumb proposed by Hurtado-Ferror et al. (2014) for Mohn's $\rho$ values is that for long-lived species, the $\rho$ value should fall between -0.15 and 0.20. 

```{r}
SShcbias(retroSummary)
```

