---
title: "Retrospective Analysis"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(ss3diags)
```

Retrospective analysis is a widely used diagnostic to test for patterns and bias in model estimated quantities (e.g. spawning stock biomass (SSB) and fishing mortality (F)). To implement a retrospective analysis with stock synthesis the `r4ss` package provides the `retro()` function. Here we provide a step-by-step example of how to run and analyze a retrospective analysis using a simple, example SS model.   

## Model inputs  
To run a stock synthesis model, 4 input files are required: starter, forecast, control, and data. The input files for the example model can be found within the `r4ss` package and accessed as shown below. If you do not have `r4ss` installed, you will need to install for this tutorial. 

```{r}
install.packages("pak")
pak::pkg_install("r4ss/r4ss")
```

```{r}
library(r4ss) 

r4ss_path <- system.file("extdata", "simple_small", package = "r4ss")
dir_retro <- file.path(tempdir(check = TRUE), "retrospectives")
dir.create(dir_retro, showWarnings = FALSE)
files <- c("data.ss", "control.ss", "forecast.ss", "starter.ss")
file.copy(from = paste(r4ss_path, files, sep = "/"), 
          to = paste(dir_retro, files, sep = "/"))

```

You will need to make sure you have the SS executable [file](https://github.com/nmfs-stock-synthesis/stock-synthesis) either in your path or in the directory you are running the retrospective from (in this case `dir_retro`). 


Once you have the 4 input files and SS executable, you can run retrospective analysis as shown below. We are running it for 5 1-year peels, so with each run, the last year of data is removed and the model is re-run for a total of 5 times. The number of year peels can be adjusted with the `years` argument. If the SS executable file you are using is named something other than "ss" (e.g. ss_opt_win.exe), you will need to specify this with the argument `exe = "ss_opt_win"`. 

```{r }

r4ss::retro(dir = dir_retro, exe = "ss", years = 0:-5)

```

To visualize the output and inspect for any patterns or biases, you need to load the report files into R and can use the `SSplotRetro()` function from `ss3diags`. The easiest way to load mulitple report files is using `r4ss::SSgetoutput()` and `r4ss::SSsummarize()` functions. Use the `keyvec` argument to indicate which folders to access, in this case since we ran 5 peels, we want the folders `retro0` to `retro-5`.

```{r}

retro_mods <- r4ss::SSgetoutput(dirvec = dir_retro, keyvec = seq(0, -5, by = -1))
retroSummary <- r4ss::SSsummarize(retro_mods)

```

